First and foremost, I have to thank Javier Marenco, my advisor. His support and guidance was invaluable in the writing of this thesis. For teaching me not just about polyhedral analysis but about academic life in general, thank you. I'd been postponing writing this thesis for quite a long time, afraid of the magnitude of the task, but you've made it a very comfortable, gradual process.

I must also thank Min Chih Lin and Isabel Méndez-Díaz, for graciously accepting to read over this thesis and provide comments.

Non academically, I thank my mom, Olga Mazzara. For always being there for me, for giving me what she had and what she didn't, for putting up with me through all these years, neither this thesis nor, in general, I, would be here. If I do not dedicate more words of gratitude to you it is because no words would be enough.

To the University of Buenos Aires. For giving me an education of the highest caliber, for allowing me to step on the same floor as giants, free of charge. For allowing me to teach here all these years, and help sustain me through my studies. For helping me reach my hopes and aspirations.

For guiding a curious, hyperactive child, and helping him find a path of his own, I thank you, Eugenio Fasano and German Rehfeldt.

Last, but most definitely not least, I would not have made it this far, nor would I be the man I am today, without the help of my friends. So to, in no particular order, Martín Miguel, Ivan Postolski, Paula Chocrón, Matías Incem, Leandro Lera Romero, Ariel Camblor, Nicolás Maffongelli, Pablo Gauna, Juan Manuel Martinez Caamaño, Pablo Fromer, Manuel Ferrería, Julián Sackmann, Gastón Bengolea Monzón, Juan Pablo Darago, Pablo Barenbaum, Saveliy Vassiliev, Xavier Warnes, Antonella Manilla, Patricio Clark, Carla Pacciarini, Rodrigo del Castillo, Maximiliano Valle, Diego Montero, Ayelén Galarza Rial, Yamila Alen, Rafael Martín, Martín Mansilla, Francisco Kordon, Karina Finkelstein, Sofía D'Alesio, Marcelo Paredes, Facundo Poggi, Pedro Tamaroff, Leandro Lovisolo, Guillermo Tochi, Fernando Martín, Luis Vion, and the friends I've definitely missed but nonetheless mean much to me, I say: Gracias. This thesis is dedicated to you.
